Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-08 manuscript.Rmd
kappa

```{r metric-stable, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}

@@

metrics_lookup <- metrics_long %>% group_by(Model, metric) %>% 
  get_summary_stats(value) %>% 
  select(-(variable:n), -(q1:mad), -se, -ci) %>%
  pivot_wider(names_from = Model, values_from = c("min","max","median", "mean", "sd"))
    column_to_rownames("metric")
```

@@

```{r metricsbyclass, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics_class.csv")}
metrics_class <- read_csv("data/group_split_metrics_class.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

Figure \@ref(fig:metrics)) and Table \@ref(tab:metrics-table)) summarize the performance of group and individual models using standard metrics for classification.

>>>>

```{r metrics-table, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics.csv")}

@@

metrics_lookup <- metrics_long %>% group_by(Model, metric) %>% 
  get_summary_stats(value) %>% 
  select(-(variable:n), -(q1:mad), -se, -ci) %>%
  pivot_wider(names_from = Model, values_from = c("min","max","median", "mean", "sd")) %>% 
    column_to_rownames("metric")
```

@@

```{r metrics-by-class, cache=TRUE, cache.extra=tools::md5sum("data/group_split_metrics_class.csv")}
metrics_class <- read_csv("data/group_split_metrics_class.csv") %>% 
  filter(model %in% c("group", "split")) %>% 
  mutate(Model = factor(model, levels = c("group", "split"), labels = c("Group", "Individual")))

@@

<!-- GET RID OF BALANCED ACCURACY AND F1, MOVE KAPPA UP, THEN DO SENS/PPV -->
Figure \@ref(fig:metrics) and Table \@ref(tab:metrics-table) summarize the performance of group and individual models using standard metrics for classification. Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Median overall accuracy was `r metrics_lookup["Overall Accuracy", "median_Group"]` for group models and `r metrics_lookup["Overall Accuracy", "median_Individual"]` for individual models. Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders. Still, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (`r metrics_lookup["Kappa", "median_Group"]`) somewhat worse compared with median individual kappas (`r metrics_lookup["Kappa", "median_Individual"]`). Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

### 2023-03-08 manuscript.tex
kappa

In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 35 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019). Accompanying our paper, we have created a database {[}CITE{]} containing synchronized video and IMU data, allowing researchers to annotate and train models for behaviors beyond body position.

@@

In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from 22 infants aged 4-14 months. Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position. A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter. Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter. The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes.

@@

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording. However, as described below, this created two video files separated by a variable gap of approximately 40 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).

@@

\hypertarget{goal-1-assess-the-proximal-accuracy-of-body-position-classification-models}{%
\subsection{Goal 1: Assess the proximal accuracy of body position classification models}\label{goal-1-assess-the-proximal-accuracy-of-body-position-classification-models}}

@@

\caption{\label{tab:metricstable}Summary statistics for model performance metrics shown separately for group and individual models.}

@@

\caption{\label{tab:metricsbyclass}Model performance metrics for each body position category, shown separately for group and individual models.}

@@

@@

@@

>>>>

In this paper, we present an inertial sensing method to classify infant body position from moment-to-moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 35 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019).

@@

In the current study, we report the feasibility and validity of body position classification over the full day in the home based on 35 testing sessions from 22 infants aged 4-14 months. Participants received a custom pair of infant leggings embedded with 4 IMUs (one on each ankle and one on each hip) and a video camera to collect ground truth data about infant body position. A \emph{proximal comparison} period began when participants received the equipment and completed a guided phone call during which caregivers were asked to elicit different body positions based on prompts from the experimenter. Although not directly supervised, this period was most similar to previous recordings because it occurred during a convenient time for the infant and caregiver to play while they received instructions from the experimenter. The \textbf{first goal} of the current study was to determine the accuracy of body position classification during the proximal comparison period using this novel, semi-supervised procedure in participants' homes. Past work found varying better performance using ``individual models''---models that were trained on one participant's data to predict their later behavior---compared with a ``group model'' that aggregated data from all infants to create a single body position classifier (Franchak et al., 2021), so we compared both modeling approaches in the current investigation.

@@

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording. However, as described below, this created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes).

@@

We report four sets of results based on 34 full-day testing sessions resulting in a total of 301.74 hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated on the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is crucial, because it provides the first ever test of how well body classification models predict behavior beyond the initial, supervised period. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day should be valid. In the third set of results we examine the data quality of full-day recordings, and in the fourth set of results we test whether full-day recordings capture well-established age differences in body position experiences.

\hypertarget{assess-the-proximal-accuracy-of-body-position-classification-models}{%
\subsection{Assess the Proximal Accuracy of Body Position Classification Models}\label{assess-the-proximal-accuracy-of-body-position-classification-models}}

@@

\caption{\label{tab:metrics-table}Summary statistics for model performance metrics shown separately for group and individual models.}

@@

\caption{\label{tab:metrics-by-class}Model performance metrics for each body position category, shown separately for group and individual models.}

@@

The first set of analyses use data from the proximal comparison, in which video annotations and motion data were tightly synchronized (within 30-60 ms). This high degree of synchronization makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work (Franchak et al., 2021), we compared two types of models. To assess the accuracy of body position classification for each recording session, we held back the last 25\% of a session's proximal comparison data as the testing set. The testing set was never used as training data. A \emph{group model} was trained for each session using an aggregated dataset containing the first 75\% of all \textbf{other} sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in training. In contrast, \emph{individual models} were trained using the first 75\% of a recording session, then tested against the last 25\%. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches.

Figure \ref{fig:metrics} and Table \ref{tab:metrics-table} summarize the performance of group and individual models using standard metrics for classification. Overall accuracy (Figure \ref{fig:metrics}A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Median overall accuracy was 0.89 for group models and 0.93 for individual models. Although individual model accuracy was greater than group model accuracy, group model accuracy approached the level of agreement found between two human coders (median = .92) indicating that group models could adequately replace human coders. Still, it is notable that the worst-case accuracy from group models (minimum = 0.50) was considerably lower than in individual models (minimum = 0.73). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent longer sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \ref{fig:metrics}), and provide classification metrics for each individual body position (Table \ref{tab:metrics-by-class}) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with median group kappas (0.77) somewhat worse compared with median individual kappas (0.85). Guidelines for interpreting kappa statistics (Landis \& Koch, 1977) consider 0.81--1.00 ``Almost Perfect,'' 0.61--0.80 ``Substantial,'' 0.41--0.60 ``Moderate,'' 0.21-- 0.40 ``Fair,'' and 0--0.20 ``Slight to Poor'', indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

@@

The start time ranged from 08:55 to 13:20 with a median of 10:20. Although the equipment was always dropped off in the morning,

@@

\leavevmode\vadjust pre{\hypertarget{ref-LandisKoch1977}{}}%
Landis, J. R., \& Koch, G. G. (1977). The measurement of observer agreement for categorical data. \emph{Biometrics}, \emph{33}(1), 159. \url{https://doi.org/10.2307/2529310}
